---
title: "conditionals_loops"
format: html
---
# Create an object called pm2_5 with a value of 48 (representing Particulate Matter 2.5, an indicator for air quality, in 
# Write an if - else if - else statement that returns “Low to moderate risk” if pm2_5 (for Particulate Matter 2.5) is less than 100, “Unhealthy for sensitive groups” if PM 2.5 is 100 <= pm2_5 < 150, and “Health risk present” if PM 2.5 is >= 150.

# Test by changing the value of your pm2_5 object and re-running your statement to check.
```{r}
pm2_5 <- 48
if (pm2_5 < 100){
  print("Low to moderate risk")
} else if (pm2_5 <= 100 & pm2_5 <150){
  print("Unhealthy for sensitive groups")
} else if (pm2_5 >= 150){
  print("Health risk present")
}
```
# Task 2
Store the string “blue whale” as an object called species. Write an if statement that returns “You found a whale!” 
```{r}
species <- c("blue whale")
if (stringr::str_detect(species, "whale")){
  print(" You found a whale!!")
}else {
  print("nothing")
}
```
# Task 3
Store the base price of a burrito as base_burrito with a value of 6.50. Store main_ingredient with a starting string of “veggie.” Write a statement that will return the price of a burrito based on what a user specifies as “main_ingredient” (either “veggie”, “chicken” or “steak”) given the following:

A veggie burrito is the cost of a base burrito
A chicken burrito costs 3.00 more than a base burrito
A steak burrito costs 3.25 more than a base burrito
```{r}
base_burrito <- 6.50
main_ingredient <- "steak"

if (main_ingredient == "veggie") {
  price = base_burrito
} else if (main_ingredient == "chicken") {
  price = base_burrito + 3.00
} else if (main_ingredient == "steak") {
  price = base_burrito + 3.25
}

print(paste("The burrito is",price, "for steak"))
```

## For Loops
# Task 4
- Create a new vector called fish that contains the values 8, 10, 12, 23 representing counts of different fish types in a fish tank (goldfish, tetras, guppies, and mollies, respectively). 
- Write a for loop that iterates through fish, and returns what proportion of total fish in the tank are that species. Assume that these counts represent all fish in the tank.
```{r}
# Add fish and = # for iteration to print out
fish <- c(goldfish = 8, tetras = 10, guppies = 12, mollies = 23)
#fish_types <- c("goldfish", "tetras", "guppies", "mollies") #not needed

for(i in seq_along(fish)){
  #Storing values into variable
  total_fish <- fish[i]/ sum(fish) #fish[i] means it will iterate over the vector. Then it will divide by the sum of the vector
  print(total_fish)
}

```
##Task 5
There is an existing vector in R called month.name that contains all month names (just try running month.name in the Console to check it out). Write a for loop that iterates over all months in month.name and prints “January is month 1,” “February is month 2”, etc.

Hint: you can index values in the month.name vector just like you would any other vector (e.g., try running month.name[5]).

```{r}
#month.name #calling data stored in R.
#month.name[5] # month of MAY

#seq_along will iterate over each month
for (i in seq_along(month.name)){
#use the dataframe with [i] for it to print the iteration and a message that says month. Lastly add another "i" so the number iteration is at the end.
print(paste(month.name[i], "is month", i))
}

```




